Paper Title*\\ {\footnotesize \textsuperscript{*}Note: Sub-titles are not captured in Xplore and should not be used} \thanks{Identify applicable funding agency here. If none, delete this.}

### 2019-05-01 icdm/submission/includes/background.tex
added example table and rescaled images

\section{Background}\label{chap:background}
This subsection gives an overview of the SGM and related work on its optimization. Furthermore, the used optimizers in our experiments are explained.
\subsection{The Skip-Gram Model}
The SGM is a very simple model used to learn WE. The idea is to train the model, a neural network, on a fake task and then use some of the weights of the neural network as embeddings. To understand this fake task center and context words must be defined. The center word is any given word in a sentence, from which we want to learn the WE. The context words of this specific center word are the words left and right in a given window $m$ in the sentence. See Figure \ref{fig:window_ex} as an example, where the context words are highlighted. An important thing to notice, is that instead of having a fixed window size $m$, each word will randomly have a window size between $1$ and $m$. The idea behind having different window sizes is that words that are further away in the sentence, have less semantic correlation to the center word.

\begin{figure}[h]
\centering
\includegraphics[scale=0.3]{images/window_ex}
\caption{Example of a center word and it's context words}
\label{fig:window_ex}
\end{figure}
%TODO think about example matrix
With these definitions the fake task can be defined: given a pair of center and context words, which will be our training samples, the goal of the network is to predict the probability of each word to appear in the context of the center word. One may ask how such a network is built. The network consists of two matrices, a projection and an output layer. Both of these matrices store word embeddings. The first one, the projection layer, will store the WE for one specific word in one specific row. It will have dimension $T\times d$, with $T$ being the size of our vocabulary and $d$ the dimension of our WE. The second layer, the output layer, will store one WE in each column. The idea behind these 2 layers is that the input layer will represent our words as center word and the output layer as context words. Since the task of the network being to predict the probability of each word appearing in the context of a given center word, the following probability should be maximized:\\
\begin{equation} \label{eq:basicSkip}
\prod_{t=1}^T \prod_{-m<j<m} p(w_{t+j}|w_t)
\end{equation}
Where T is the number of words in the corpus data, $w_t$ the $t^{th}$ word in the corpus data and $m$ is the context window. This means that the $m$ nearest words to $w$ are considered as context words.
Equation \ref{eq:basicSkip} can be transformed into sums by using log probabilities:
\begin{equation}
\sum _{t=1}^T \sum_{-m<j<m} log( p(w_{t+j}|w_t) )
\end{equation}
where the parameters are the same as in Equation \ref{eq:basicSkip}.

@@

\begin{align*}
p(c|w) &= p(y=1|c,w) + \prod_{k\in K} p(y=0|k,c)
\\&= p(y=1|c,w) + \prod_{k\in K} 1- p(y=1|k,c)
\\&= log((p(y=1|c,w)) + \sum_{k\in K} log(1- p(y=1|k,c))
\\&= log(\frac{1}{1+e^{-v_c \tilde{v_w }}}) + \sum_{k\in K} log(1-\frac{1}{1+e^{-v_c \tilde{v_k}}})
\\&= log(\frac{1}{1+e^{-v_c \tilde{v_w } }}) + \sum_{k\in K} log(\frac{1}{1+e^{v_c \tilde{v_k} }})
\\&= log(\sigma(v_c \tilde{v_w } ) + \sum_{k\in K} \sigma(log(-v_c \tilde{v_k} )) &&\text{where, $\sigma(x) = \frac{1}{1+e^{-x}}$}
\end{align*}\label{eq:obj_neg_samples}

Where $v_c$ and $\tilde{v_w }$, can be interpreted as before in Equation \ref{eq:softmax}. The goal is to maximize this objective function. Another way to look at this function, apart from logistic regression, and understand why it's working so well, is to assume that two vectors are similar if their dot product is high. Therefore, we maximize the dot product of similar words $(w,c)$ ($c$ appears in the context of $w$) and minimize it for dissimilar words $(w,k_i)$ (those were selected randomly).
We see that to compute our objective function we will only have to compute the sum over the number of negative samples $K$, which is very small in practice (2-20). To put things in perspective lets imagine our data set consists of 100000 words, we set $K=2$. Assume each output neuron has a weight vector $v$ with $|v| = 300$. In consequence, when updating our weights we would only update $0.2*10^{-2}$ of the 300 million weights in the output layer.

@@

\begin{equation} \label{eq:unigram}
P(w)=\frac{f(w)^{\frac{3}{4}}}{\sum_{t=0}^{T} f(w_t)^{\frac{3}{4}}}
\end{equation}
where $f(w)$ is the frequency of $w_t$. The value of $\frac{3}{4}$ is set empirically. Raising the unigram distribution to power of $\frac{3}{4}$ makes it less likely for a word to be drawn if it often appears in the dataset in comparison to the basic unigram distribution. See figure \ref{fig:frequency_ex} for an example.
\begin{figure}[ht]
\centering
\includegraphics[scale=0.45]{images/frequency_ex}
\caption{Probabililty of a word, of the text8 dataset (sampled), to be chosen at random according to its frequency and the power to which the unigram distribution is raised}
\label{fig:frequency_ex}
\end{figure}
It's quite easily observable that this approach will outperform the classical softmax in computation time, as it only compute the sum over $K$ output nodes. Now the question arises if the accuracy is good enough but according to Mikolov et al. \cite{mikolov2} the negative sampling method "is an extremely simple training method that learns accurate representations". As a matter of fact, Mikolov et al. \cite{mikolov2} reported a 6\% accuracy improvement in comparison to a Hierarchical Softmax model. Therefore, it is a good solution to the problem raised by the classical softmax.
We now have enough background knowledge about the SGM to look at how it can be optimized. The next subsection introduces earlier approaches to optimize the SGM.

>>>>

\section{Background}\label{chap:background}

\begin{table*}[tb]\centering
    \caption{Sample Caption}
    %\scriptsize
    \begin{tabular}{l r r r r r r r}%
        \toprule
        \textbf{App} & \textbf{NSGA-II} & \textbf{Random} & \textbf{Heuristic} & \textbf{Simple {ga}} & \textbf{MOSA} & \textbf{MIO} \\%
        \midrule%
        moneywallet        & 0.20 & 0.40 & 0.40 & 0.22 & 0.22 & 0.19 \\%
        shoppinglist       & 0.77 & 0.78 & 0.84 & 0.77 & 0.77 & 0.77 \\%
        periodical         & 0.85 & 0.86 & 0.87 & 0.84 & 0.83 & 0.85 \\%
        drhoffmannsoftware & 0.44 & 0.48 & 0.45 & 0.44 & 0.46 & 0.48 \\%
        activitydiary      & 0.74 & 0.82 & 0.82 & 0.73 & 0.74 & 0.75 \\%
        bierverkostung     & 0.43 & 0.65 & 0.63 & 0.44 & 0.45 & 0.49 \\%
        easy\_xkcd         & 0.62 & 0.73 & 0.68 & 0.57 & 0.59 & 0.60 \\%
        markor             & 0.62 & 0.68 & 0.65 & 0.61 & 0.60 & 0.60 \\%
        redreader          & 0.55 & 0.61 & 0.50 & 0.52 & 0.54 & 0.52 \\%
        rentalcalc         & 0.67 & 0.77 & 0.76 & 0.68 & 0.68 & 0.71 \\%
        \midrule%
        Mean               & 0.59 & 0.68 & 0.66 & 0.58 & 0.59 & 0.60 \\%
        \bottomrule%
   \end{tabular}%
   \label{tab:sample_label}%
\end{table*}
This subsection gives an overview of the SGM and related work on its optimization. Furthermore, the used optimizers in our experiments are explained.
\subsection{The Skip-Gram Model}
The SGM is a very simple model used to learn WE. The idea is to train the model, a neural network, on a fake task and then use some of the weights of the neural network as embeddings. To understand this fake task center and context words must be defined. The center word is any given word in a sentence, from which we want to learn the WE. The context words of this specific center word are the words left and right in a given window $m$ in the sentence. See Figure \ref{fig:window_ex} as an example, where the context words are highlighted. An important thing to notice, is that instead of having a fixed window size $m$, each word will randomly have a window size between $1$ and $m$. The idea behind having different window sizes is that words that are further away in the sentence, have less semantic correlation to the center word.

\begin{figure}[h]
\centering
\includegraphics[scale=0.18]{images/window_ex}
\caption{Example of a center word and it's context words}
\label{fig:window_ex}
\end{figure}
%TODO think about example matrix
With these definitions the fake task can be defined: given a pair of center and context words, which will be our training samples, the goal of the network is to predict the probability of each word to appear in the context of the center word. One may ask how such a network is built. The network consists of two matrices, a projection and an output layer. Both of these matrices store word embeddings. The first one, the projection layer, will store the WE for one specific word in one specific row. It will have dimension $T\times d$, with $T$ being the size of our vocabulary and $d$ the dimension of our WE. The second layer, the output layer, will store one WE in each column. The idea behind these 2 layers is that the input layer will represent our words as center word and the output layer as context words. Since the task of the network being to predict the probability of each word appearing in the context of a given center word, the following probability should be maximized:\\
\begin{equation} \label{eq:basicSkip}
\prod_{t=1}^T \prod_{-m<j<m} p(w_{t+j}|w_t)
\end{equation}
Where T is the number of words in the corpus data, $w_t$ the $t^{th}$ word in the corpus data and $m$ is the context window. This means that the $m$ nearest words to $w$ are considered as context words.
Equation \ref{eq:basicSkip} can be transformed into sums by using log probabilities:
\begin{equation}
\sum _{t=1}^T \sum_{-m<j<m} log( p(w_{t+j}|w_t) )
\end{equation}
where the parameters are the same as in Equation \ref{eq:basicSkip}.

@@

\begin{align*}
p(c|w) &= p(y=1|c,w) + \prod_{k\in K} p(y=0|k,c)
\\&= p(y=1|c,w) + \prod_{k\in K} 1- p(y=1|k,c)
\\&= log((p(y=1|c,w)) + \sum_{k\in K} log(1- p(y=1|k,c))
\\&= log(\frac{1}{1+e^{-v_c \tilde{v_w }}}) + \sum_{k\in K} log(1-\frac{1}{1+e^{-v_c \tilde{v_k}}})
\\&= log(\frac{1}{1+e^{-v_c \tilde{v_w } }}) + \sum_{k\in K} log(\frac{1}{1+e^{v_c \tilde{v_k} }})
\\&= log(\sigma(v_c \tilde{v_w } ) + \sum_{k\in K} \sigma(log(-v_c \tilde{v_k} ))
\end{align*}\label{eq:obj_neg_samples}

Where $v_c$ and $\tilde{v_w }$, can be interpreted as before in Equation \ref{eq:softmax} and $\sigma(x) = \frac{1}{1+e^{-x}}$. The goal is to maximize this objective function. Another way to look at this function, apart from logistic regression, and understand why it's working so well, is to assume that two vectors are similar if their dot product is high. Therefore, we maximize the dot product of similar words $(w,c)$ ($c$ appears in the context of $w$) and minimize it for dissimilar words $(w,k_i)$ (those were selected randomly).
We see that to compute our objective function we will only have to compute the sum over the number of negative samples $K$, which is very small in practice (2-20). To put things in perspective lets imagine our data set consists of 100000 words, we set $K=2$. Assume each output neuron has a weight vector $v$ with $|v| = 300$. In consequence, when updating our weights we would only update $0.2*10^{-2}$ of the 300 million weights in the output layer.

@@

\begin{equation} \label{eq:unigram}
P(w)=\frac{f(w)^{\frac{3}{4}}}{\sum_{t=0}^{T} f(w_t)^{\frac{3}{4}}}
\end{equation}
where $f(w)$ is the frequency of $w_t$. The value of $\frac{3}{4}$ is set empirically. Raising the unigram distribution to power of $\frac{3}{4}$ makes it less likely for a word to be drawn if it often appears in the dataset in comparison to the basic unigram distribution. See figure \ref{fig:frequency_ex} for an example.
\begin{figure}[ht]
\centering
\includegraphics[scale=0.30]{images/frequency_ex}
\caption{Probabililty of a word, of the text8 dataset (sampled), to be chosen at random according to its frequency and the power to which the unigram distribution is raised}
\label{fig:frequency_ex}
\end{figure}
It's quite easily observable that this approach will outperform the classical softmax in computation time, as it only compute the sum over $K$ output nodes. Now the question arises if the accuracy is good enough but according to Mikolov et al. \cite{mikolov2} the negative sampling method "is an extremely simple training method that learns accurate representations". As a matter of fact, Mikolov et al. \cite{mikolov2} reported a 6\% accuracy improvement in comparison to a Hierarchical Softmax model. Therefore, it is a good solution to the problem raised by the classical softmax.
We now have enough background knowledge about the SGM to look at how it can be optimized. The next subsection introduces earlier approaches to optimize the SGM.
